### 2020-10-27 submission/manuscript.tex
Edit Results section

\author{}
\date{\vspace{-2.5em}}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{setspace}
\doublespacing
\usepackage[left]{lineno}
\linenumbers
\modulolinenumbers
\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
\usepackage[T1]{fontenc}

@@

\subsection{word choice: sample
vs.~community}\label{word-choice-sample-vs.community}

@@

@@

\textbf{\emph{Results.}}

\textbf{Datasets.} I collected 12 publicly available datasets that used
the Illumina MiSeq platform to sequence the V4 region of the 16S rRNA
gene from a variety of environments (Table 1). To insure the highest
possible data quality, datasets were limited to those where the 500
cycle v2 MiSeq chemistry was used to sequence the amplicons. The paired
250 nt reads resulted in near complete 2-fold sequencing coverage of
every nucleotide in the ca. 250 nt-long region. This region and
sequencing platform were selected because previous work has shown that a
standard data analysis pipeline in mothur results in a sequencing error
rate below 0.02\% (\textbf{???}). All sequence data were obtained from
the Sequence Read Archive and processed using a standard mothur-based
sequencing pipeline that resulted in ASVs as generated by the
pre.cluster algorithm (\textbf{???}, 1). After removing poor quality and
chimeric ASVs and samples that had uncharacteristically low number of
sequences for the dataset, these datasets included between 7.00 and
490.00 samples. The median number of sequences for each dataset ranged
between 6,477.00 and 193,464.00 (Table 1, Figure S1). Strikingly, aside
from the relatively small marine and soil datasets, the difference
between the sample with the fewest sequences and the sample with the
most sequences for each dataset varied between 7.4 and 96.6-fold (Table
1).

\textbf{The nature of singletons.} Removal of rare ASVs is commonly
justified as a method of removing ASVs that are artifacts. If such ASVs
are artifacts, then one would expect the number of singleton ASVs to
accumulate with sequencing depth. Contrary to this expectation, the
median percentage of sequences that were discarded when singleton ASVs
were removed from each dataset varied between 0.42 and 22.23\%
(bioethanol and seagrass). In addition, with the exception of the
samples from the marine and sediment datasets (Spearman correlation,
P\textgreater{}0.05), the fraction of singleton ASVs in the sample was
negatively correlated with the number of sequences in each sample with a
range between -0.27 and -0.87 (rice and bioethanol). This showed that
with additional sequencing, the probability of seeing rare ASVs multiple
times was greater than the probability of generating an artifact. This
suggests that the rare ASVs are not as likely to be artifacts as
previously thought. Furthermore, if rare ASVs were artifacts, then one
would not expect to find them in other samples from the same dataset. In
fact, singleton ASVs from samples with fewer sequences were often found
in samples with more sequences. At least 50\% of the singleton ASVs
found in the samples from the mice, rice, seagrass, and stream datasets
were found in another sample from the same dataset. Considering the
likelihood of finding an ASV duplicated in another sample is confounded
by the number of samples and inter-sample diversity, the high coverage
of singleton ASVs in these datasets was remarkable. The correlation
between the number of sequences in a sample and the fraction of that
sample's singleton ASVs that were covered by another sample in the
dataset was significant and negative for 9 of the datasets ranging
between -0.31 and -0.84 for the rice and seagrass datasets,
respectively. The negative correlation indicated that the singleton ASVs
in the smaller samples were more likely to be covered by ASVs in the
larger samples. Among the three datasets without a significant
correlation (Spearman correlation, P\textgreater{}0.05), the marine and
soil datasets had the fewest samples in our collection and the stream
dataset already had a high level of coverage regardless of the number of
sequences. Contrary to the common motivation for removing rare ASVs,
these results indicate that this practice disproportionately impacts
samples with fewer sequences and likely removes a large number of
non-artifact ASVs.

\textbf{How do communities change when rare ASVs are removed?} Removing
rare ASVs will reduce the number of observed ASVs and proportionally
increase the relative abundance of the remaining ASVs. The result was
expected to be a loss of information contained within each sample. To
quantify the effect of removing rare ASVs on the information contained
within each sample, I varied the minimum abundance threshold to simulate
removing ASVs of varying rarity from each sample. The number of ASVs in
each sample decreased by between 34.4 and 86.2\% when removing those
ASVs that only appeared once and by between 76.0 and 95.6\% when
removing those that appeared ten or fewer times from each sample (Figure
1). Similarly, the Shannon diversity decreased by between 1.8 and 15.9\%
when removing ASVs that only appeared once and by between 76.0 and
95.6\% when removing ASVs that appeared ten or fewer times from each
sample (Figure 1). Next, I assigned the ASVs to OTUs to assess the
impact of removing rare ASVs on higher level taxonomic groupings that
are commonly used in microbial ecology studies. Although pooling similar
ASVs into OTUs reduced the impact of removing the rare ASVs relative to
the ASV-based analysis, the minimum abundance threshold still decreased
the number of observed OTUs and the diversity decreased relative to the
original community (\textbf{Figure 1}). In contrast to the number of
ASVs or OTUs and their diversity, the Kullback--Leibler divergence
compares the relative abundance of each ASV or OTU between
representations of the community. I calculated the Kullback--Leibler
divergence from the full to pruned communities when rare ASVs were
removed. As the threshold for removing ASVs increased, the amount of
information lost also increased for both ASVs and OTUs (Figure 1). The
relative loss of information was generally lower for OTUs than than it
was for ASVs. Removing rare ASVs, regardless of abundance threshold, had
profound impacts on the representation of the communities.

\textbf{Removing treatment group effects from community data.} Because
treatment effects often affect a sample's diversity and inter-sample
variation, I generated null distributions for each study by randomizing
the number of times each ASV was observed in each sample such that the
total number of sequences in each sample and the total number of times
each ASV was observed across all samples in the study was the same as
was originally observed. This effectively made every community in a
study a statistical sample of the study-wide composite community
distribution. For example, after this procedure, the 490 samples from
the human dataset would be expected to have the same number of ASVs and
diversity and one would not expect to find treatment-based effects
between the samples. Because of the risk of bias if only one
representation of the null distribution was generated, I generated 100
randomized datasets for each study. The same trends between removing
rare ASVs and the number of ASVs, sample diversity, and information loss
that were identified using the observed community community distribution
data were also identified with the data from the null distribution
(\textbf{Figure S2}). The null distribution data were used in the
remainder of the study.

\textbf{How do the relationships between communities change when they
are screened by removing rare ASVs.} Considering the loss of the number
of ASVs, diversity, and information when a community has its rarest ASVs
removed, it seemed likely that the relationship between communities
would be altered. To assess the impact of removing rare ASVs on measures
of alpha diversity between samples I calculated the coefficients of
variation (COV, i.e.~standard deviation divided by the mean) for the
number of ASVs and diversity for each study at multiple abundance
thresholds. The COV for the number of ASVs across the studies after
removing singletons were between 3.6 and 32.7-times larger than they
were without removing singleton ASVs. Similarly, the COVs for the
diversity of ASVs were between 1.8 and 20.4-times larger when singletons
were removed than when they were not removed.To assess the impact of
removing rare ASVs on measures of beta diversity between samples I
calculated the COV of the Bray-Curtis distances between samples within
the same study at multiple abundance thresholds. The COV between
Bray-Curtis distances within a study when singletons were removed was
between 1.3 and 18.6-times larger than when they were not removed.
Increasing the minimum abundance threshold increased the COV between
samples when using metrics of alpha and beta diversity. When ASVs were
clustered into OTUs the difference in the COV was less than it was for
the ASVs (\textbf{Figure 2}). These results indicate that removing rare
ASVs increases the dissimilarity between samples, which could have a
significant impact on the statistical power to detect differences
between treatment groups.

\textbf{How does the increased inter-sample variation impact the ability
to detect differences between samples?} To test the effect of increased
inter-sample variation, I randomly assigned samples to one of two
treatment groups. In the first treatment group, communities were
randomly sampled from the null distribution as described above. For the
second treatment group, I increased the abundance of 10\% of the ASVs in
the pooled study distribution by 5\%. I randomly generated 100 simulated
sets of treatment groups and samples. I then tested the ability to
detect a difference between the two treatment groups using alpha and
beta diversity metrics. The fraction of significant tests was a
measurement of the statistical power to detect the difference between
the treatment groups. Across all of the metrics I tested, the marine
dataset yielded no simulated sets that were statistically significant,
which was likely due to the small number of samples in the study (N=7);
the marine dataset is not included in the following analysis. Among the
remaining datasets, the power to detect a difference in the observed
number of ASVs ranged between 0.10 and 0.49 and between 0.10 and 0.53 to
detect a difference in diversity when using a Wilcox test
(\textbf{Figure X}). When singleton ASVs were removed, the power to
detect a difference in the number of ASVs dropped by between 27.3 and
92.9\% and by between 40.0 and 93.3\% to detect a difference in
diversity. The effect of removing rare ASVs on the number of OTUs and
their diversity was similar. A popular alternative approach calculates
the dissimilarity between communities based on the abundance of each ASV
or OTU. We used the Bray-Curtis dissimilarity index to compare the
simulated communities within each dataset and calculated the power to
detect differences between the two simulated treatment groups using the
analysis of molecular variance (also called PERMANOVA). Without removing
rare sequences, the power to detect a difference between the two
simulated treatment groups varied between 0.41 and 1.00. Aside from 3
datasets, the power to detect differences dropped by between 6.5 and
64.0\% when singletons were removed. However, when ASVs that occurred 10
or fewer times were removed from each sample, the power to detect
differences dropped by 12.0 and 97.2\%; similar results were observed
when ASVs were clustered into OTUs. Removing rare ASVs reduced the
ability to detect simulated treatment effects using metrics commonly
used to compare microbial communities.

\textbf{How does the removing of rare ASVs impact the probability of
falsely detecting a difference between treatment groups?} Observing
reduced ability to detect differences between communities when rare ASVs
were removed from each sample, I next asked whether removing rare ASVs
could lead to falsely claiming that a treatment effect had a significant
effect on community diversity and structure. First, for each dataset, I
sampled sequences from the null distribution and randomly assigned each
sample to one of two treatment groups and determined the number of
observed ASVs and OTUs and their Shannon diversity index. Testing at an
experiment-wise error rate of 0.05, I expected 5\% of the iterations for
each dataset to yield a significant test result. Indeed, there was no
evidence that removing rare ASVs resulted in an inflated experiment-wise
error rate. The average fraction of significant tests did not
meaningfully vary from 0.05 across the minimum abundance threshold,
dataset, metric of describing sample alpha-diversity, or whether the
abundance of ASVs or OTUs were used (\textbf{Figure XA}). Similarly, the
average fraction of significant tests did not meaningfully vary from
0.05 when using analysis of molecular variance to compare communities
using Bray-Curtis distances. Second, I again sampled sequences from the
null distribution, but assigned samples to one of two treatment groups
based on the number of sequences in each sample. The samples with fewer
than the median number of sequences for the dataset were assigned to one
group and those with more than the median were assigned to the other.
This exaggerated bias has been observed in comparisons of the lung and
oral microbiota because of the larger nubmer of non-specific amplicons
that can be sequenced from lung samples relative to those in the oral
cavity {[}\textbf{REF}{]}. When rare sequences were not removed, the
fraction of significant tests did not differ from 5\% for comparing the
number of observed taxa (using ASVs or OTUs), their shannon diversity,
or Bray-Curtis distances. However, when rare taxa of any frequency were
removed, the probability of falsely detecing a difference as signifiant
increased with the definition of rarity (\textbf{Figure XB}). Across the
datasets, the average fraction the average fraction of falsely detecting
a difference was when only singletons were removed was greater than
90\%. If there is any relationship between the number of sequences and
the treatment group, the risk of falsely rejecting the null hypothesis
is inflated when researchers use the strategy of removing rare
sequences. The most conservative approach is to not remove low abundance
sequences.

@@

1. \textbf{Schloss PD}, \textbf{Westcott SL}, \textbf{Ryabin T},
\textbf{Hall JR}, \textbf{Hartmann M}, \textbf{Hollister EB},
\textbf{Lesniewski RA}, \textbf{Oakley BB}, \textbf{Parks DH},
\textbf{Robinson CJ}, \textbf{Sahl JW}, \textbf{Stres B},
\textbf{Thallinger GG}, \textbf{Horn DJV}, \textbf{Weber CF}. 2009.
Introducing mothur: Open-source, platform-independent,
community-supported software for describing and comparing microbial
communities. Applied and Environmental Microbiology
\textbf{75}:7537--7541.
doi:\href{http://dx.doi.org/10.1128/aem.01541-09}{10.1128/aem.01541-09}.

>>>>

\author{}
\date{\vspace{-2.5em}}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{setspace}
\doublespacing
\usepackage[left]{lineno}
\linenumbers
\modulolinenumbers
\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
\usepackage[T1]{fontenc}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}

@@

@@

\textbf{word choice: sample vs.~community}

@@

\subsection{Results}\label{results}

\textbf{Datasets.} I collected 12 publicly available datasets that used
the Illumina MiSeq platform to sequence the V4 region of the 16S rRNA
gene from a variety of environments (Table 1). To insure the highest
possible data quality, datasets were limited to those where the 500
cycle v2 MiSeq chemistry was used to sequence the amplicons. The paired
250 nt reads resulted in near complete 2-fold sequencing coverage of
every nucleotide in the ca. 250 nt-long region. This region and
sequencing platform were selected because previous work has shown that a
standard data analysis pipeline in mothur results in a sequencing error
rate below 0.02\% (1). All sequence data were obtained from the Sequence
Read Archive and processed using a standard mothur-based sequencing
pipeline that resulted in ASVs as generated by the pre.cluster algorithm
(1, 2). After removing poor quality and chimeric ASVs and samples that
had uncharacteristically low number of sequences for the dataset, these
datasets included between 7 and 490 samples (Figure S1). The median
number of sequences for each dataset ranged between 6,477 and 193,464
(Table 1). Strikingly, aside from the relatively small marine and soil
datasets, the difference between the sample with the fewest sequences
and the sample with the most sequences for each dataset varied by
between 7.4 and 96.6-fold (Table 1).

\textbf{The nature of singletons.} Removal of rare ASVs is commonly
justified as a method of removing ASVs that are artifacts. If such ASVs
are artifacts, then one would expect the number of singleton ASVs to
accumulate with sequencing depth. Contrary to this expectation, the
median percentage of sequences that were discarded when singleton ASVs
were removed from each dataset varied between 0.42 and 22.23\%
(bioethanol and seagrass). In addition, with the exception of the
samples from the marine and sediment datasets (Spearman correlation,
P\textgreater{}0.05), the fraction of singleton ASVs in samples was
negatively correlated with the number of sequences in each sample with a
range between -0.27 and -0.87 (rice and bioethanol) (Figure 1A). This
showed that with additional sequencing, the probability of seeing
singleton ASVs in multiple samples was greater than the probability of
generating an artifact. This suggests that the singleton ASVs are not as
likely to be artifacts as previously thought. Furthermore, if singleton
ASVs were artifacts, then one would not expect to find them in other
samples from the same dataset. In fact, singleton ASVs from samples with
fewer sequences were often found in samples with more sequences. At
least 50\% of the singleton ASVs found in the samples from the mice,
rice, seagrass, and stream datasets were found in another sample from
the same dataset (Figure 1B). Considering the likelihood of finding an
ASV duplicated in another sample is confounded by the number of samples
and inter-sample diversity, the high coverage of singleton ASVs in these
datasets was remarkable. The correlation between the number of sequences
in a sample and the fraction of that sample's singleton ASVs that were
covered by another sample in the dataset was significant and negative
for 9 of the datasets ranging between -0.31 and -0.84 for the rice and
seagrass datasets, respectively (Figure 1C). The negative correlation
indicated that the singleton ASVs in the smaller samples were more
likely to be covered by ASVs in the larger samples. Among the three
datasets without a significant correlation (Spearman correlation,
P\textgreater{}0.05), the marine and soil datasets had the fewest
samples in our collection and the stream dataset already had a high
level of coverage regardless of the number of sequences. Contrary to the
common motivation for removing rare ASVs, these results indicate that
this practice disproportionately impacts samples with fewer sequences
and likely removes more non-artifact ASVs than those that are artifacts.

\textbf{The impact of removing rare ASVs on the information represented
in each sample.} Removing rare ASVs will reduce the richness of ASVs and
proportionally increase the relative abundance of the remaining ASVs.
The result was expected to be a loss of information contained within
each sample. To quantify the effect of removing rare ASVs on the
information contained within each sample, I varied the minimum abundance
threshold to simulate removing ASVs of varying rarity from each sample.
The richness of ASVs in each sample (i.e.~the number of ASVs) decreased
by between 34.4 and 86.2\% when removing those ASVs that only appeared
once and by between 76.0 and 95.6\% when removing those that appeared
ten or fewer times from each sample (Figure 2A). Similarly, the Shannon
diversity decreased by between 1.8 and 15.9\% when removing ASVs that
only appeared once and by between 76.0 and 95.6\% when removing ASVs
that appeared ten or fewer times from each sample (Figure 2B). Next, I
assigned the ASVs to OTUs to assess the impact of removing rare ASVs on
higher level taxonomic groupings that are commonly used in microbial
ecology studies. Although pooling similar ASVs into OTUs reduced the
impact of removing the rare ASVs relative to the ASV-based analysis, the
minimum abundance threshold still decreased the richness of OTUs and the
diversity decreased relative to the full community (\textbf{Figure
S2AB}). In contrast to the richness and diversity measurements, the
Kullback--Leibler divergence compares the relative abundance of specific
ASVs or OTUs between representations of the community. I calculated the
Kullback--Leibler divergence from the full to pruned communities when
rare ASVs were removed. As the threshold for removing ASVs increased,
the amount of information lost also increased for both ASVs and OTUs
(Figure 2C and Figure S2C). The relative loss of information was
generally lower for OTUs than than it was for ASVs. Removing rare ASVs,
regardless of abundance threshold, had profound impacts on the
representation of the communities.

\textbf{Removing treatment group effects from community data.} Because
treatment effects often affect a sample's diversity and inter-sample
variation, I generated null distributions for each study by randomizing
the number of times each ASV was observed in each sample such that the
total number of sequences in each sample and the total number of times
each ASV was observed across all samples in the study was the same as
was originally observed. This effectively made every community in a
study a statistical sample of the study-wide composite community
distribution. For example, after this procedure, the 490 samples from
the human dataset would be expected to have the same richness and
diversity of ASVs and one would not expect to find treatment-based
effects between the samples. Because of the risk of bias if only one
representation of the null distribution was generated, I generated 100
randomized datasets for each study. The trends between removing rare
ASVs and the richness, diversity, and information loss that were
identified using the observed community community distribution data were
also identified with the data from the null distribution; however, the
losses were larger when using the null distribution data (\textbf{Figure
S3}). The null distribution data were used in the remainder of the study
to minimize the risk of bias.

\textbf{The impact of removing rare ASVs on the information represented
between samples.} Considering the loss of richness, diversity, and
information when a community has its rarest ASVs removed, it seemed
likely that the relationship between communities would also be altered.
To assess the impact of removing rare ASVs on measures of alpha
diversity between samples I calculated the coefficients of variation
(COVs, i.e.~standard deviation divided by the mean) for richness and
diversity for each study at multiple abundance thresholds. The COVs for
richness of ASVs across the studies after removing singletons were
between 3.6 and 32.7-times larger than they were without removing
singleton ASVs (Figure 3A). Similarly, the COVs for the diversity of
ASVs were between 1.8 and 20.4-times larger when singletons were removed
than when they were not removed (Figure 3B). To assess the impact of
removing rare ASVs on measures of beta diversity between samples I
calculated the COVs of the Bray-Curtis distances between samples within
the same study at multiple abundance thresholds. The COVs between
Bray-Curtis distances within a study when singletons were removed was
between 1.3 and 18.6-times larger than when they were not removed
(Figure 3C). Increasing the minimum abundance threshold increased the
COVs between samples when using metrics of alpha and beta diversity.
When ASVs were clustered into OTUs the difference in COVs was less than
it was for the ASVs (Figure S4). These results indicate that removing
rare ASVs increases the dissimilarity between samples, which could have
a significant impact on the statistical power to detect differences
between treatment groups.

\textbf{The impact of removing rare ASVs on the ability to detect
statistically significant differences between treatment groups.} To test
the effect of increased inter-sample variation, I randomly assigned
samples to one of two treatment groups. In the first treatment group,
communities were randomly sampled from the null distribution as
described above. For the second treatment group, I increased the
abundance of 10\% of the ASVs in the pooled study distribution by 5\%. I
randomly generated 100 simulated sets of treatment groups and samples. I
then tested the ability to detect a difference between the two treatment
groups using alpha and beta diversity metrics. The fraction of
significant tests was a measurement of the statistical power to detect
the difference between the treatment groups. When considering the
differences in richness and diversity, the marine dataset yielded no
simulated sets that were statistically significant, which was likely due
to the small number of samples in the study (N=7). Among the remaining
datasets, the power to detect a difference in the richness of ASVs
ranged between 0.10 and 0.49 and between 0.10 and 0.53 to detect a
difference in diversity when using a Wilcox test (Figure 4A). When
singleton ASVs were removed, the power to detect a difference in the
diversity of ASVs dropped by between 27.3 and 92.9\% and by between 40.0
and 93.3\% (Figure 4B). The effect of removing rare ASVs on the richness
of OTUs and their diversity was similar (Figure S5AB). I used the
Bray-Curtis dissimilarity index to compare the simulated communities
within each dataset and calculated the power to detect differences
between the two simulated treatment groups using the analysis of
molecular variance (also called PERMANOVA) (Figure 4C and S5C). Without
removing rare sequences, the power to detect a difference between the
two simulated treatment groups varied between 0.41 and 1.00. Aside from
3 datasets, the power to detect differences dropped by between 6.5 and
64.0\% when singletons were removed. However, when ASVs that occurred 10
or fewer times were removed from each sample, the power to detect
differences dropped by 12.0 and 97.2\%; similar results were observed
when ASVs were clustered into OTUs. Removing rare ASVs reduced the
ability to detect simulated treatment effects using metrics commonly
used to compare microbial communities.

\textbf{The impact of removing rare ASVs on the probability of falsely
detecting a difference between treatment groups.} Observing reduced
ability to detect differences between communities when rare ASVs were
removed from each sample, I next asked whether removing rare ASVs could
lead to falsely claiming that a treatment effect had a significant
effect on community diversity and structure. First, I sampled sequences
from the null distribution for each dataset and randomly assigned each
sample to one of two treatment groups and determined the richness and
diversity of ASVs and OTUs. Testing at an experiment-wise error rate of
0.05, I expected 5\% of the iterations for each dataset to yield a
significant test result. Indeed, there was no evidence that removing
rare ASVs resulted in an inflated experiment-wise error rate. The
average fraction of significant tests did not meaningfully vary from
0.05 across the minimum abundance threshold, dataset, metric of
describing sample alpha-diversity, or whether the abundance of ASVs or
OTUs were used (Figure 5A and S6A). Similarly, the average fraction of
significant tests did not meaningfully vary from 0.05 when using
analysis of molecular variance to compare communities using Bray-Curtis
distances (Figure 5A and S6A). Second, I again sampled sequences from
the null distribution, but assigned samples to one of two treatment
groups based on the number of sequences in each sample. The samples with
fewer than the median number of sequences for the dataset were assigned
to one group and those with more than the median were assigned to the
other. This exaggerated bias has been observed in comparisons of the
lung and oral microbiota because of the larger number of non-specific
amplicons that can be sequenced from lung samples relative to those in
the oral cavity leading to a significant difference in sequencing depth
between treatment groups {[}\textbf{REF}{]}. When rare sequences were
not removed, the fraction of significant tests did not differ from 5\%
for comparing the richness, their diversity, or Bray-Curtis distances
(Figure 5B and S6B). However, when rare taxa of any frequency were
removed, the probability of falsely detecing a difference as signifiant
increased with the definition of rarity (Figure 5B and S6B). Not
including the small marine dataset, the average fraction the average
fraction of falsely detecting a difference across datasets when only
singletons were removed was 92.45\%. If there is any relationship
between the number of sequences and the treatment group, the risk of
falsely rejecting the null hypothesis is inflated when researchers use
the strategy of removing rare sequences. The most conservative approach
is to not remove low abundance sequences.

@@

1. \textbf{Kozich JJ}, \textbf{Westcott SL}, \textbf{Baxter NT},
\textbf{Highlander SK}, \textbf{Schloss PD}. 2013. Development of a
dual-index sequencing strategy and curation pipeline for analyzing
amplicon sequence data on the MiSeq Illumina sequencing platform.
Applied and environmental microbiology \textbf{79}:5112--5120.

2. \textbf{Schloss PD}, \textbf{Westcott SL}, \textbf{Ryabin T},
\textbf{Hall JR}, \textbf{Hartmann M}, \textbf{Hollister EB},
\textbf{Lesniewski RA}, \textbf{Oakley BB}, \textbf{Parks DH},
\textbf{Robinson CJ}, \textbf{Sahl JW}, \textbf{Stres B},
\textbf{Thallinger GG}, \textbf{Horn DJV}, \textbf{Weber CF}. 2009.
Introducing mothur: Open-source, platform-independent,
community-supported software for describing and comparing microbial
communities. Applied and Environmental Microbiology
\textbf{75}:7537--7541.
doi:\href{http://dx.doi.org/10.1128/aem.01541-09}{10.1128/aem.01541-09}.
